---
title: "Parameterized ggplots"
author: "Michael Ingrisch"
date: "March 31, 2017"
output: 
  html_document: 
    keep_md: yes
---
## Problem

I was always wondering how exactly ggplot resolves the column names that you provide e.g. in aesthetic mappings.
This became a problem in the moment in which I wanted to wrap a ggplot call into a function, in order to quickly plot several columns of a complex dataset. Here, I reproduce this issue in a simple example.
We are working in the tidyverse and use the built-in mpg dataframe to keep things simple:
```{r setup, echo=T, message = F, warning = F}
library(tidyverse)
head(mpg)

```
This is a very simple function that wraps around a ggplot call and returns a plot:
```{r plot.data.function}
plot.data <- function(data){
  g <- ggplot(data, aes(x = displ, y=hwy))+
    geom_point()
  g
}
plot.data(mpg)
```
Problems begin to arise once we try to pass a variable that is supposed to determine the y aesthetics - e.g. cty:

```{r plot.data.function.param}
plot.data.param <- function(data, param){
  g <- ggplot(data, aes(x = displ, y=param))+
    geom_point()
  g
}
plot.data.param(mpg, 'cty')
```
This error makes sense, since `cty`is not known when our function is called. Of course, we can pass the _name_ of the column as a string - but then, ggplot would try to map a single value, the string, on the y axis:

```{r}
plot.data.param(mpg, 'cty')
```

## The solution: aes_string

Therefore, we need to find a way to pass a string variable to the ggplot aesthetics. The solution is provided by aes_string:

```{r plot.data.aes.string}
plot.data.aes.string <- function(data, param){
  str(param)
  g <- ggplot(data, aes_string(x = 'displ',y= param))+
    geom_point()
  g
}
plot.data.aes.string(mpg, 'hwy')
plot.data.aes.string(mpg, 'cty')
```
Problem solved.


